---
author: "Eric Hsienchen Chu"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      error = TRUE, fig.align="center", fig.height = 5, fig.width = 8)
library(tidyverse)
library(stringr)
library(lubridate)
#library(plm)       # Panel data analysis library
#library(car)       # Companion to applied regression 
#library(tseries)   # For timeseries analysis
#library(lmtest)    # For hetoroskedasticity analysis
#library(directlabels)
#library(ggrepel)
#library(ggtext)
#library(gganimate)
#library(gifski)
#library(kableExtra)
#library(ggpubr)
#library(rddtools)
#library(magrittr)
#library(maps)
#library(mapdata)
library(broom)  # Convert models to data frames
#library(rdrobust)  # For robust nonparametric regression discontinuity
#library(estimatr)  # Run 2SLS models in one step with iv_robust()
#library(ivreg)
#library(ivmodel)
library(gtools)

library(sna)
library(igraph)
library(threejs)


library(sna)
library(igraph)
library(threejs)

library(data.table)
library(dplyr)
library(readstata13)


#source("../viridis.R")
```

## [ECON664: Gravity Models] 

#### Author: Eric Hsienchen Chu

#### Date: September 27, 2023


- - -

#### Part I. Data Cleaning

```{r}
## At the top of the file, always include rm(list = ls()) to clear memory. 
rm(list = ls())
```

```{r} 
## Read in the data. Since this is a .dta file, we use the read.dta13
## command. If this were a .csv or .excel file, you would use fread,
## read.table, or one of the read excel packages. 

## Check working directory
data <- read.dta13("../ECON 664/data_exercise_class/data/class_gravity1.dta")
```


```{r}
## head(data) shows you what the top of the data frame looks like. 
head(data)
```


```{r}
## Create Variables for "Strict" version of the gravity regression. 
data = data %>% mutate(
                     lvalue  = log(value),
                     lgdp_od = log(gdp_o*gdp_d),
                     ldistw  = log(distw),
                     lgdp_o  = log(gdp_o),
                     lgdp_d  = log(gdp_d)
                 ) 

data = data %>% mutate(lindex = lgdp_od - ldistw)
```



```{r}
## Plot the data in 1950 to see what it looks like
data_1950 = data %>% filter(year == 1950)

pdf('index_scatter_1950.pdf', height = 5, width = 5)
plot(data_1950$lindex, data_1950$lvalue, 
     col = 'blue', 
     ylim = c(3,15), 
     pch = 20, 
     xlim = c(15,35),
     xlab = 'Gravity Index', 
     ylab = 'Log Export Value (Origin to Destination)')
text(data_1950$lindex, data_1950$lvalue-1, labels=data_1950$iso_od, cex = 0.3)
dev.off()
```


- - - 

#### Part II. Gravity Model Regressions

#### Regression 1 (Strict Gravity)


```{r}
###################################
## Regression 1 (Strict Gravity) ##
###################################

strict_gravity_reg = lm(lvalue~lindex, data = data %>% filter(year == 1950))

## summary(reg) will print out the regression table
summary(strict_gravity_reg)

tidy(strict_gravity_reg) %>% 
  mutate(sig = stars.pval(p.value))
```


#### Regression 2 (General Gravity)


```{r}
####################################
## Regression 2 (General Gravity) ##
####################################

general_gravity_reg = lm(lvalue~lgdp_o + lgdp_d + ldistw, data = data %>% filter(year == 1950))

general_gravity_estimates = data.frame(summary(general_gravity_reg)$coefficients)

general_gravity_estimates
```

`Note.` the coeff estimates of lgdp_o is 0.6726462

> Interpretation: Holding other variables constant, in the year 1950, a 1% increase in origin's GDP raises bilateral trade by 0.67%. 

```{r}
general_gravity_reg_2019 = lm(lvalue~lgdp_o + lgdp_d + ldistw, data = data %>% filter(year == 2019))

general_gravity_estimates_2019 = data.frame(summary(general_gravity_reg_2019)$coefficients)

general_gravity_estimates_2019
```

> If we run the general gravity on 2019 data, we observe that the signs are very stable.

- - -

#### Part III. Has the world gotten flatter?

> Does "distance" matter LESS? We can examine how the distance coefficient has changed over time!

```{r}
##################################
## Has the world gotten flatter?##
##################################

## Create variable stripping market size effects out of trade flows (controling for market size)
## First, we have to get coefficients from the last regression

b_cons   = general_gravity_estimates$Estimate[rownames(general_gravity_estimates)=='(Intercept)']
b_lgdp_o = general_gravity_estimates$Estimate[rownames(general_gravity_estimates)=='lgdp_o']
b_lgdp_d = general_gravity_estimates$Estimate[rownames(general_gravity_estimates)=='lgdp_d']
b_ldistw = general_gravity_estimates$Estimate[rownames(general_gravity_estimates)=='ldistw']

## Now, we'll use these to create the new variable, lvalue_hat
data <- data %>% mutate(lvalue_hat = lvalue - (b_cons + b_lgdp_o*lgdp_o + b_lgdp_d*lgdp_d))
```



```{r}
## If we run a regression of lvalue_hat on ldsitw in 1950, we should get the same coefficient as before
## It looks good. 
lm(lvalue_hat~ldistw, data = data %>% filter(year == 1950))
```


```{r}
## Now, we'll write a loop to run this regression in each year.

ylist <- c(1950,1960,1970,1980,1990,2000,2010,2019)
coeff <- data.frame() ## Here, I initialized an empty data frame to collect the coefficients.

for (yy in ylist) {

    reg  = lm(lvalue_hat~ldistw, data = data %>% filter(year == yy))
    reg  = data.frame(summary(reg)$coefficients)

    temp = data.frame(
        year = yy,
        beta = reg$Estimate[rownames(reg)=='ldistw'],
        se   = reg$Std..Error[rownames(reg)=='ldistw'])
                   
    coeff = rbind(coeff, temp)
}
```


```{r}
## This is some code to make a plot of these coefficients

## First, we can calculate the 95 percent confidence intervals
coeff = coeff %>% mutate(                       
                       lower95 = (beta - se*qnorm(1-(1-.95)/2)),
                       upper95 = (beta + se*qnorm(1-(1-.95)/2))
                   )

pdf(paste0('distance_coeff_over_time.pdf'), height = 5, width = 5)
plot(coeff$year, coeff$beta, pch = 17, ylim = c(-1.5,-0.4), col = 'steelblue', xlab = 'Year', ylab = 'Coefficient on Log Distance')
arrows(x0=coeff$year, y0=coeff$lower95, x1 = coeff$year, y1 = coeff$upper95, code = 3, angle=90, length=0.05, col = 'gray')
points(coeff$year, coeff$beta, pch = 17, col = 'steelblue')
```

> So, does the world get "flatter"? NO! (the coeff on log-distance is increasing in absolute value (more negative))

> Consider f(x,t) = 0 and f(x, t') = the curve of the plot, then f(x, t')-f(x, t) ≥ 0 -> supermodular.


```{r}
#dev.off()
```

> Takeawys: Trade flows respond *positively* to market size. Trade flows respond *negatively* to distance. The relationship is incredibly stable and is one of the most robust empirical results in (international trade) economics. 

---

**Eric Hsienchen Chu**

UW-Madison Economics





